# Code-a-thon Project Template

This project is intended to help you apply your skills in R in a way that would be conducive for collaborating or creating reproducible analyses.


## Data Used

Identify a dataset to analyze. Remember that if you share the data you must **not violate any privacy restrictions**

If you don't have your own data, options for places to find data are listed on https://daseh.org.

You might start by setting up an R Project, importing your data, previewing the data, etc. 


```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(dygraphs)
library(xts)          # To make the convertion data-frame / xts format
library(lubridate)

```

```{r}
read_csv("Gakona_USEPA_PM25_AQI.csv")
read_csv("Galena_USEPA_PM25_AQI.csv")
read_csv("PortAlsworth_USEPA_PM25_AQI.csv")
read_csv("Talkeena_USEPA_PM25_AQI.csv")
```



**describe** what your data looks like and **where** you got it. Identify what the variables and samples are, or point to a data dictionary. Describe how the data was originally created. 

This is air quality index data from two PurpleAir monitors (low-cost air quality sensors) in rural Alaska, specifically Gakona and Louden AK. The data is 8640 10-minute average AQI observations from two sensors (A and B) for each monitor, between May 30th 2024 to July 29th 2024. 


# Data Cleaning/Wrangling

Perform any data subsetting, cleaning, or manipulation methods that were described in this course on your data. Examples are: renaming the columns, recoding values, reshaping the data, filtering the data etc. 

You might also research other methods to include here, especially if your data is very unique.

```{r}
data2sens <- select(Galena_USEPA_PM25_AQI, "DateTime", "GAKONA, AK A", "GAKONA, AK B", "Louden Tribal Office A", "Louden Tribal Office B")
```

```{r}
data2sens <- rename(data2sens, Gakona_A = `GAKONA, AK A`)
```

```{r}
data2sens <- rename(data2sens, Gakona_B = `GAKONA, AK B`)
```

```{r}
data2sens <- rename(data2sens, Louden_A = "Louden Tribal Office A")
data2sens <- rename(data2sens, Louden_B = "Louden Tribal Office B")
```

```{r}
data2sens <- mutate(data2sens, datetime=as.POSIXct(DateTime, tz="", format="%Y-%m-%d %H:%M:%S"))

```
#Trying to make a new column with average of A and B sensors for each location.
```{r}
#New column that takes the average AQI values of the Gakona sensors A and B

data2sens$Gakona_avg <- (data2sens$Gakona_A + data2sens$Gakona_B)/2
```

```{r}
#New column that takes the average AQI values of the Louden sensors A and B

data2sens$Louden_avg <- (data2sens$Louden_A + data2sens$Louden_B)/2
```

```{r}
GLavg <- select(data2sens, "DateTime", "Gakona_avg", "Louden_avg")
```

```{r}
#Restructure data table to change colors of lines. Create new column with 
GLavg <- 
```


# Data Visualization

Create some visualizations of your data using the esquisse app or the ggplot2 package.

```{r}
#Plot comparing Gakona sensors A and B
data2sens_plot <-
  ggplot() + 
  geom_line(data2sens, mapping = aes(x = DateTime, y = Gakona_A, color="A")) +
  geom_line(data2sens, mapping = aes(x = DateTime, y = Gakona_B, color="B")) +
  ggtitle("Gakona, AK PM2.5 AQI 10min Avg - 5/30/24 to 7/29/24") +
  xlab("Date") + ylab("PM2.5 AQI") 

data2sens_plot
```

```{r}
#Plot comparing Louden sensors A and B
data2sens_plot2 <-
  ggplot() + 
  geom_line(data2sens, mapping = aes(x = DateTime, y = Louden_A, color="A")) +
  geom_line(data2sens, mapping = aes(x = DateTime, y = Louden_B, color="B")) +
  ggtitle("Louden, AK PM2.5 AQI 10min Avg - 5/30/24 to 7/29/24") +
  xlab("Date") + ylab("PM2.5 AQI") 

data2sens_plot2
```

```{r}
#Plot comparing the Gakona AQI averages and Louden AQI averages
data2sens_plot3 <-
  ggplot() + 
  geom_line(data2sens, mapping = aes(x = DateTime, y = Gakona_avg, color="Gakona")) + 
  geom_line(data2sens, mapping = aes(x = DateTime, y = Louden_avg, color="Louden")) + 
  annotate(geom="point", x=as.POSIXct("2024-06-30 22:40:00"), y=153.5, 
             size=5, shape=21, fill = "transparent") +
  ggtitle("Gakona vs. Louden, AK PM2.5 AQI 10min Avg - 5/30/24 to 7/29/24") +
  xlab("Date") + ylab("AQI") +
  annotate("rect", xmin = as.POSIXct("2024-05-30 12:10:00"), xmax = as.POSIXct("2024-07-29 12:00:00"), ymin = -1, ymax = 50, alpha = .2, fill = "green") +
  annotate("rect", xmin = as.POSIXct("2024-05-30 12:10:00"), xmax = as.POSIXct("2024-07-29 12:00:00"), ymin = 50, ymax = 100, alpha = .2, fill = "yellow") +
  annotate("rect", xmin = as.POSIXct("2024-05-30 12:10:00"), xmax = as.POSIXct("2024-07-29 12:00:00"), ymin = 100, ymax = 150, alpha = .2, fill = "orange") +
  annotate("rect", xmin = as.POSIXct("2024-05-30 12:10:00"), xmax = as.POSIXct("2024-07-29 12:00:00"), ymin = 150, ymax = 165, alpha = .2, fill = "red") +
   theme(
    legend.position = c(.95, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(3, 3, 3, 3)
    ) +
  guides(fill=guide_legend(title="Sensor Location"))

data2sens_plot3
```

```{r}
#Maximum value
filter(data2sens, Louden_avg == 153.5)
```


```{r}
# Interactive plot for Gakona_A data.
# Then you can create the xts necessary to use dygraph
don <- xts(x = data2sens$Gakona_A, order.by = data2sens$datetime)
```

```{r}
# Finally the plot
p <- dygraph(don) %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors="#D8AE5A") %>%
  dyRangeSelector() %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
  dyRoller(rollPeriod = 1)
```


```{r}
# Interactive plot (dygraph) comparing Gakona avg and Louden avg sensor values. 
# Create the xts necessary to use dygraph
don2 <- xts(x = data2sens$Gakona_avg, order.by = data2sens$datetime)
```

```{r}
# Finally the plot
p2 <- dygraph(don2) %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors="#D8AE5A") %>%
  dyRangeSelector() %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 5, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
  dyRoller(rollPeriod = 1)
```




# Data Analysis

Perform a **simple analysis** of your data. This can involve summarizing the data to describe aspects about it (quartiles, means, range etc.) or a simple statistical test. 

Don't forget to describe what analysis you performed and why. Provide some simple **interpretation** about what your analysis might indicate about your data.

Want to see how many days 

```{r}

```


# Please include additional sections as desired.

```{r}

```


# Versions

Include some information about the version of R and the packages you are using by running sessionInfo().

```{r}
sessionInfo()
```

Render your document to an html or pdf using the "Knit" button.
